Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-23 writeup/gz_hubble_data.tex

The dimming and resolution effects of redshift can significantly affect galaxy
classifications derived from any method. To facilitate corrections due to
redshift, we include a sample of 288 galaxies with SDSS imaging that can be
transformed to simulate \hst{} imaging out to $z = 1$. The selection spanned a
variety of galaxy morphologies (as selected by GZ2 classifications) and
$r^\prime$-band surface brightnesses, and also spanned the redshift range of
SDSS targets (in $N_z = 4$ bins) in order to optimize for different target
minimum redshifts in \hst{} imaging. We redshifted the galaxies and produced
simulated \hst{} imaging using the \ferengi{} code \citep{bar08a}.

%The selection criteria for the different morphological categories is
%summarized in Table \ref{tbl:morphologies}. 
The surface brightness selection ($N_\mu = 3$) for the \ferengi{} galaxies was
(1) low: $\mu > 21.5$~\magarc;  (2) mid: $20.5 < \mu < 21.5$~\magarc; and (3)
high: $\mu < 20.5$~\magarc. For each of the four ``target redshifts''
($z_\mathrm{sim} = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in
$\Delta z = 0.1$ bins up to $z_\mathrm{sim}=1.0$.

@@

In addition to the physical parameters of the input images, the \ferengi{}
output depends on assumptions of the global galaxy evolution model. This
evolution is a crude mechanism that mimics the brightness increase of galaxies
with increasing redshift (out to at least $z\sim1-2$). The effect on the
redshifted images is simply an empirical addition to the magnitude of a galaxy
of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and
$e$ is the evolutionary correction in magnitudes (i.e., $e=-1$ essentially
brightens the galaxy by 1~magnitude by $z=1$). \ferengi{} was run on the images
for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments
of $\Delta e = 0.5$. Figure~\ref{fig:exampleFERENGI} shows several examples
of the effects of ``losing'' spiral/disc features with increasing redshift
for two galaxies with $e=0$.

The final number of \ferengi{} images produced for each galaxy is ultimately a
function of galaxy's redshift (since the new images cannot be resampled at
better angular resolution than the original SDSS data), as well as the number
of $e$ values selected. Table~\ref{tbl:ferengivalues} summarizes the redshifted
images produced for GZH. The use of these images to correct for redshift-dependent
bias is described in Section \ref{sec:debiasing}.

\begin{table}
\caption{Summary of \ferengi{} artificial redshifting \label{tbl:ferengivalues}}
\begin{tabular}{lccccr}
\hline\hline
$z_{\rm sim}$ & $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm galaxies}$ & $N_{\rm images}$\\
\hline
0.3              & 8                  & 7                   & $-3.0$        & 72             & 4032 \\
0.5              & 6                  & 4                   & $-1.5$        & 72             & 1728 \\
0.8              & 3                  & 3                   & $-1.0$        & 72             &  648 \\
1.0              & 1                  & 3                   & $-1.0$        & 72             &  216 \\
\hline\hline
\end{tabular}
\end{table}

@@

Classifications were aggregated and weighted only if the classifier was logged
into GZH under their username (which was encouraged, but not required for
participation). Classifications by participants who were not logged in were marked as
``Anonymous'' and assumed to have $w = 1$. 
% We should also try to estimate how many anonymous classifiers there were, if possible.

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin11,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS) which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the morphological distribution
were due to a redshift-dependent bias that affects image quality and
classification accuracy (ie, galaxies of a given mass/size will appear smaller
and dimmer at higher redshifts).  This bias is not unique to the crowdsourcing
technique; its dependence in data quality is a potential problem for
\emph{both} automated and visual classifications, and must be addressed in
order to accurately measure demographics over any significant redshift range.

@@

In the GZH samples, the redshift range is large enough that cosmological
evolution of the types and morphologies of galaxies is expected for galaxies in
the sample. As a result, the previous methods of correcting for
redshift-dependent bias do not work. In addition, the effects of band shifting
will change the images even more across these redshift ranges. 
%Figure~\ref{fig:exampleFERENGI} illustrates some of the possible effects of
%losing features in spiral galaxies at high redshift.

In order to test and correct for the effects of redshift, GZH includes a set of
calibration images. These images consist of the same galaxy as it would appear
over a variety of redshifts. The input images are from the SDSS
\citep{yor00,str02} and are processed to match the observational properties of
the \hst{} surveys out to $z=1$. These images were classified in the GZH
interface using the same classification scheme as the original \hst{} images.
Further details on the creation of these images is given in Section
\ref{ssec:ferengi}.

The GZH classifications include 288 unique galaxies originally generated from
SDSS imaging and processed with the \ferengi{} code. The selection spanned a
variety of galaxy morphologies (as selected by GZ2 classifications) and
$r^\prime$-band surface brightnesses, and also spanned the redshift range of
SDSS targets that maximize the coverage of parameter space
(Figure~\ref{fig:sb_redshift}).

@@

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. There may also be other parameters that affect
classifier accuracy, such as the proximity of close companions or 
individual classifier ability and skill. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters intended to capture the bulk of the change in bias: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

This dependence of the vote fraction on $\mu_r$ and \zsim{} is measured using the
\ferengi{} images over all the evolutionary correction factors. It is assumed
that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy can be expressed as:

@@

\noindent where $f_{\mu,z=0.3}$ is the ``calibrated'' vote fraction at the
lowest redshift in the \ferengi{} bins ($z=0.3$) and $\zeta$ is a positive
parameter that controls the rate at which $f$ decreases with increasing
redshift. This formula fits the data relatively well (with very few exceptions,
the vote fractions for featured galaxies decrease monotonically with increasing
redshift), and the exponential function bounds the observed vote fractions
between $f_{\mu,z=0.3}$ and zero. Figure~\ref{fig:zeta_examples} shows the
change in vote fraction and the fit to Equation~\ref{eqn:fzeta} for a random
selection of galaxies in the \ferengi{} images.

@@

GZH uses the values of $\zeta$ for \emph{all} sets of artificially redshifted
galaxies to fit the overall distribution as a function of surface brightness,
since the correction being applied is expected to vary as a function of the
intrinsic galaxy properties. The galaxies that can be used to measure the
calibration are restricted to those with a non-zero $f_\textrm{features}$ at
$z_\mathrm{sim}=0.3$ and with a reasonable fit ($\Delta \chi^2 > 3.0$) to the
exponential model for the measured change in \ffeatures.

Figure~\ref{fig:zeta_mu} shows the results of fitting the \ferengi{} images
with Equation~\ref{eqn:fzeta}; the correction is only a weak function of
surface brightness ($\mu$). Higher-surface brightness galaxies have stronger
average corrections, likely because these galaxies are more likely to have
larger $f_\textrm{features}$ values at high redshifts. Low surface brightness
galaxies are more likely to begin low and remain low; the bounded nature of the
dropoff (and variance among the individual voters) means that the average
magnitude of $\zeta$ will be lower.

\begin{figure}
\center
\includegraphics[width=0.50\textwidth]{figures/zeta_mu.png}
\caption{All fits for the vote fraction dropoff parameter $\zeta$ for
\ffeatures{} in the \ferengi{} galaxies as a function of surface
brightness. This includes only those galaxies with a reasonably bounded range
on the dropoff ($-10<\log(\zeta)<10$) and sufficient points to fit each
function (37 original galaxies, each artificially redshifted to create a total
of 296~images}
\label{fig:zeta_mu}
\end{figure}

The data in Figure~\ref{fig:zeta_mu} are fit with a linear function such that:

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/p_vs_p_SB_redshift.pdf}
\caption{Effects of redshift bias in 3,950 images in the \ferengi{} sample.
Each point in a given redshift and surface brightness bin represents a
unique galaxy. On the $y$-axis in each bin is the \ffeatures{} value of the
image of that galaxy redshifted to the value corresponding to that redshift
bin. On the $x$-axis is the \ffeatures{} value of the image of the same galaxy
redshifted to $z=0.3$. The dashed black lines represent the best-fit
polynomials to the data in each square. The solid black line represents
\ffeaturesz=\ffeaturesrest. Regions in which there is a single-valued
relationship between \ffeatures{} at high redshift and at $z=0.3$ are white;
those in which there is not are blue, and those with not enough data ($N<5$)
are gray. A larger version of the bin outlined at $z=1.0$ and $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$ is shown in Figure~\ref{fig:f_vs_f_zoom}.}
\label{fig:f_vs_f}
\end{figure*}

@@

Figure~\ref{fig:f_vs_f} shows the change in \ffeatures{} for the \ferengi{}
galaxies relative to their lowest simulated redshift. In this analysis, only
galaxies whose lowest simulated redshift image was $z_\mathrm{sim}=0.3$ were
used (see Table \ref{tbl:ferengivalues}), and only those which had detectable
surface brightness measurements in \sextractor; this includes 3,950 of the
total 6,466 images. For each simulated redshift value \zsim{} at a fixed
surface brightness $\mu$, \ffeaturesz, the value measured at that simulated
redshift, is plotted against \ffeaturesrest, the value measured for the same
galaxy at \zsim$=0.3$.

>>>>

The dimming and resolution effects of redshift can significantly affect galaxy
classifications derived from any method. To facilitate corrections due to
redshift, we include a sample of 288 galaxies with SDSS imaging that can be
transformed to simulate \hst{} imaging out to $z = 1$. 
The SDSS images were redshifted and processed to mimic simulated \hst{} imaging 
parameters using the \ferengi{} code \citep{bar08a}.

%The selection criteria for the different morphological categories is
%summarized in Table \ref{tbl:morphologies}.

@@

In addition to the physical parameters of the input images, the \ferengi{}
output depends on assumptions of the global galaxy evolution model. This
evolution is a crude mechanism that mimics the brightness increase of galaxies
with increasing redshift (out to at least $z\sim1-2$). The effect on the
redshifted images is simply an empirical addition to the magnitude of a galaxy
of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and
$e$ is the evolutionary correction in magnitudes (i.e., $e=-1$ essentially
brightens the entire galaxy by 1~magnitude by $z=1$). \ferengi{} was run on the images
for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments
of $\Delta e = 0.5$. Figure~\ref{fig:exampleFERENGI} shows several examples
of the effects of ``losing'' spiral/disc features with increasing redshift
for two galaxies with $e=0$.

The final number of \ferengi{} images produced for each galaxy is ultimately a
function of galaxy's redshift (since the new images cannot be resampled at
better angular resolution than the original SDSS data), as well as the number
of $e$ values selected. The use of these images to correct for redshift-dependent
bias is described in Section \ref{sec:debiasing}.

%\begin{table}
%\caption{Summary of \ferengi{} artificial redshifting \label{tbl:ferengivalues}}
%\begin{tabular}{lccccr}
%\hline\hline
%$z_{\rm sim}$ & $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm galaxies}$ & $N_{\rm images}$\\
%\hline
%0.3              & 8                  & 7                   & $-3.0$        & 72             & 4032 \\
%0.5              & 6                  & 4                   & $-1.5$        & 72             & 1728 \\
%0.8              & 3                  & 3                   & $-1.0$        & 72             &  648 \\
%1.0              & 1                  & 3                   & $-1.0$        & 72             &  216 \\
%\hline\hline
%\end{tabular}
%\end{table}

@@

Classifications were aggregated and weighted only if the classifier was logged
into GZH under their username (which was encouraged, but not required for
participation). Classifications by participants who were not logged in were marked as
``Anonymous'' and receive the same $w=1$ weighting as the vast majority
of logged-in classifiers.
% We should also try to estimate how many anonymous classifiers there were, if possible. 
% Not clear that I can do that given the SQL data

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin11,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS), which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the morphological distribution
were due to a redshift-dependent bias that affects image quality and
classification accuracy (\ie, galaxies of a given mass/size will appear smaller
and dimmer at higher redshifts).  This bias is not unique to the crowdsourcing
technique; its dependence in data quality is a potential problem for
\emph{both} automated and visual classifications, and must be addressed in
order to accurately measure demographics over any significant redshift range.

@@

In GZH, the redshift range is large enough that cosmological
evolution of the types and morphologies of galaxies is expected for the \hst{}
sample. As a result, the previous methods of correcting for
redshift-dependent bias do not work. In addition, the effects of band shifting
will change the images even more across these redshift ranges.

In order to test and correct for the effects of redshift, GZH includes a set of
calibration images.  These are simulated images of a set of nearby galaxies as
they would appear observed at a variety of redshifts.  The input images are
from the SDSS \citep{yor00,str02}.  The sample consists of 288~nearby galaxies
at $z<0.013$. The galaxies spanned a variety of morphologies (as selected by
GZ2 classifications, including identifications of spiral structure,
ellipticals, mergers, edge-on disks, bulge prominence/shape, and bars) and
$r^\prime$-band surface brightnesses.  The overall selection spanned the
redshift range of SDSS targets and maximized the number of \hst{} galaxies at
the same surface brightness and redshift (Figure~\ref{fig:sb_redshift}).  The
\ferengi{} code was used to produce sets of images corresponding to
observations of these galaxies out to $z_\mathrm{sim}=1.0$ (see
Section~\ref{ssec:ferengi}).  The resulting images were classified in the GZH
interface using the standard classification scheme.

@@

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

%This dependence of the vote fraction on $\mu_r$ and \zsim{} is measured using the \ferengi{} images over all the evolutionary correction factors. 
We model the debiased vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy as:

@@

\noindent where $f_{\mu,z=0.3}$ is the vote fraction at the
lowest redshift in the \ferengi{} bins ($z=0.3$). Here, $\zeta$ is a model
parameter that controls the rate at which \ffeatures{} decreases with increasing
redshift (and that may depend on other galaxy properties). This formula fits
the data relatively well (with very few exceptions, the vote fractions for
featured galaxies decrease monotonically with increasing redshift), and the
exponential function bounds the observed vote fractions between $f_{\mu,z=0.3}$
and zero. Figure~\ref{fig:zeta_examples} shows the change in vote fraction and
the best-fit model for a random sample of galaxies.

@@

The data from Figure~\ref{fig:zeta_examples} show that there is a clear
dependence on redshift for the observed changes in \ffeatures. To examine
whether a better fit should involve different parameters, we examine the global
dependence of $\zeta$ on the galaxy surface brightness. This is motivated by
the fact that brighter galaxies would presumably be easier to identify;
\citet{bam09} and \citet{wil13} found changes in \ffeatures{} depending on a
combination of physical size and absolute magnitude. To simplify our model, we
subsume these into a single parameter $\mu$.

Figure~\ref{fig:zeta_mu} examines the results of fitting the \ferengi{} images
with Equation~\ref{eqn:fzeta}. Interestingly, the correction is only a weak
function of surface brightness. Higher-surface brightness galaxies have
slightly stronger average corrections, possibly because these galaxies are more
likely to have larger $f_\textrm{features}$ values at high redshifts. Low
surface brightness galaxies are more likely to begin low and remain low; the
bounded nature of the dropoff (and variance among the individual voters) means
that the average magnitude of $\zeta$ will be lower.

\begin{figure}
\center
\includegraphics[width=0.50\textwidth]{figures/zeta_mu.png}
\caption{All fits for the \ferengi{} galaxies of the vote fraction dropoff
parameter $\zeta$ for \ffeatures{} as a function of surface brightness. This
includes only galaxies with a bounded range on the dropoff
($-10<\log(\zeta)<10$) and sufficient points to fit each function (37~original
galaxies, each artificially redshifted to create a total of 296~images).}
\label{fig:zeta_mu}
\end{figure}

Since there is little evidence for any strong systematic dependence of $\zeta$
on $\mu$, we do not include any additional parameters in fitting to our
calibration model.  We fit the data in Figure~\ref{fig:zeta_mu} with a linear
function:

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/p_vs_p_SB_redshift.pdf}
\caption{Effects of redshift bias in 3,950 images in the \ferengi{} sample.
Each point \emph{in a given redshift and surface brightness bin} represents a
unique galaxy. On the $y$-axis in each bin is the \ffeatures{} value of the
image of that galaxy redshifted to the value corresponding to that redshift
bin. On the $x$-axis is the \ffeatures{} value of the image of the same galaxy
redshifted to $z=0.3$. The dashed black lines represent the best-fit
polynomials to the data in each square. The solid black line represents
\ffeaturesz=\ffeaturesrest. Regions in which there is a single-valued
relationship between \ffeatures{} at high redshift and at $z=0.3$ are white;
those in which there is not are blue, and those with not enough data ($N<5$)
are gray. A larger version of the bin outlined at $z=1.0$ and $20.3 < \mu <
21.0$ $\rm (mag/arcsec^2)$ is shown in Figure~\ref{fig:f_vs_f_zoom}.}
\label{fig:f_vs_f}
\end{figure*}

@@

Figure~\ref{fig:f_vs_f} shows the change in \ffeatures{} for the \ferengi{}
galaxies relative to their lowest simulated redshift. This analysis only uses
galaxies whose lowest simulated redshift image was $z_\mathrm{sim}=0.3$ and
with detectable surface brightness measurements from \sextractor; this includes
3,950 of the total 6,466~images originally processed. For each simulated
redshift value \zsim{} at a fixed surface brightness $\mu$, \ffeaturesz, the
value measured at that simulated redshift, is plotted against \ffeaturesrest,
the value measured for the same galaxy at \zsim$=0.3$.
